cardinalR: Generating interesting high-dimensional data structures

### 2025-10-13 paper-cardinalR.Rmd
Need draw the implementation diagram

```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of."}
knitr::include_graphics("figures/cardinalR.pdf")
```

@@

circle_proj1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.75, 0.8), 
  title = "a1", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

@@

```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="80%", fig.align='center', fig.pos="H"}

@@

```{r summaryclust-tb-html, eval=knitr::is_html_output()}
summary_clust |> 
  kable(caption = "cardinalR branching data generation functions", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), table.pos = "H") 
```

```{r summaryclust-tb-pdf, eval=knitr::is_latex_output()}
summary_clust |> 
  kable(caption = "cardinalR branching data generation functions", format="latex", col.names = c("Metric", "wb.ratio", "Dunn_Index", "Corrected_Rand", "VI"), booktabs = T, table.pos = "H")  |>
  column_spec(1, width = "2.5cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "2cm") |>
  column_spec(4, width = "2cm") |>
  column_spec(5, width = "2cm")
```

Overall, all methods produced similar compactness and separation, as reflected by the *within–between cluster ratios (wb.ratio)* and *Dunn indices*. However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups. In comparison, $k$-means and hierarchical clustering showed moderate agreement with the true labels. These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.

>>>>

```{r workflow, out.width = "100%", fig.pos="H", fig.cap="Workflow of.", eval=FALSE}
knitr::include_graphics("figures/cardinalR.pdf")
```

@@

circle_proj1 <- plot_proj(
  proj_obj = proj_obj1, 
  point_param = c(1.5, 0.2, "#000000"), # size, alpha, color
  plot_limits = c(-0.8, 0.8), 
  title = "a1", 
  cex = 2, 
  axis_text_size = 5,
  is_color = FALSE)

@@

```{r sphere-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `circle` (a1-a3), `curvycycle` (b1-b3), `unifsphere` (c1-c3), `gridedsphere` (d1-d3), `clusteredspheres` (e1-e3), and `hemisphere` (f1-f3) data.", fig.width=12, fig.height=12, out.width="100%", fig.align='center', fig.pos="H"}

@@

```{r summaryclust-tb-html, eval=knitr::is_html_output()}
summary_clust |> 
  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), tab.pos = "H") 
```

```{r summaryclust-tb-pdf, eval=knitr::is_latex_output()}
summary_clust |> 
  kable(caption = "Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.", format="latex", col.names = c("Metric", "wb.ratio", "Dunn Index", "Corrected Rand", "VI"), booktabs = T, table.pos = "H")  |>
  column_spec(1, width = "2.5cm") |>
  column_spec(2, width = "2cm") |>
  column_spec(3, width = "2cm") |>
  column_spec(4, width = "2.5cm") |>
  column_spec(5, width = "2cm")
```

Overall, all methods produced similar compactness and separation, as reflected by the *within–between cluster ratios (wb.ratio)* and *Dunn indices*. However, the **model-based clustering** achieved the highest *Corrected Rand Index* ($0.75$) and lowest *Variation of Information (VI)* ($0.65$), indicating the best recovery of the true underlying groups (Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:summaryclust-tb-html)', '\\@ref(tab:summaryclust-tb-pdf)'))`). In comparison, $k$-means and hierarchical clustering showed moderate agreement with the true labels. These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.

### 2025-10-13 paper-cardinalR.tex
Regenerate embeddings with new data
Need draw the implementation diagram

\begin{figure}[H]
\includegraphics[width=1\linewidth]{figures/cardinalR} \caption{Workflow of.}\label{fig:workflow}
\end{figure}

@@

{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}

@@

{\centering \includegraphics[width=0.8\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}

@@

\caption{\label{tab:summaryclust-tb-pdf}cardinalR branching data generation functions}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
\toprule
Metric & wb.ratio & Dunn\_Index & Corrected\_Rand & VI\\
\midrule
k-means & 0.61 & 0.01 & 0.42 & 1.32\\
Hierarchical & 0.61 & 0.01 & 0.50 & 1.15\\
Model-based & 0.61 & 0.01 & 0.75 & 0.65\\
\bottomrule
\end{tabular}
\end{table}

Overall, all methods produced similar compactness and separation, as reflected by the \emph{within--between cluster ratios (wb.ratio)} and \emph{Dunn indices}. However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups. In comparison, \(k\)-means and hierarchical clustering showed moderate agreement with the true labels. These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.

>>>>

@@

{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/sphere-proj-1}

@@

{\centering \includegraphics[width=1\linewidth]{paper-cardinalR_files/figure-latex/triginometric-proj-1}

@@

\caption{\label{tab:summaryclust-tb-pdf}Comparison of clustering performance metrics (within–between ratio (wb.ratio), Dunn index, Corrected Rand index, and variation of information (VI) across $k$-means, hierarchical, and model-based clustering methods.}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2.5cm}>{\raggedleft\arraybackslash}p{2cm}}
\toprule
Metric & wb.ratio & Dunn Index & Corrected Rand & VI\\
\midrule
k-means & 0.61 & 0.01 & 0.42 & 1.32\\
Hierarchical & 0.61 & 0.01 & 0.50 & 1.15\\
Model-based & 0.61 & 0.01 & 0.75 & 0.65\\
\bottomrule
\end{tabular}
\end{table}

Overall, all methods produced similar compactness and separation, as reflected by the \emph{within--between cluster ratios (wb.ratio)} and \emph{Dunn indices}. However, the \textbf{model-based clustering} achieved the highest \emph{Corrected Rand Index} (\(0.75\)) and lowest \emph{Variation of Information (VI)} (\(0.65\)), indicating the best recovery of the true underlying groups (Table \ref{tab:summaryclust-tb-pdf}). In comparison, \(k\)-means and hierarchical clustering showed moderate agreement with the true labels. These findings demonstrate that mixture-based approaches can more effectively capture the heterogeneity of clusters in high-dimensional, non-spherical data.
